---
title: "DataAnalysisLab2"
author: "Joy Chowdhury"
date: "February 3, 2017"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R

## Listing 3-0

Listing 3-0 demonstrates how the working directory can be set using R and how a vector of package names can be compared against installed packages to determine new packages to be installed, saving bandwidth and time.  The packages are used to create graphs and other visuals.

```{r 3-0}
setwd("~/Class Notes and Assignments/SRT411/SRT411-DataAnalysisLab-2/")
pkg <- c("ggplot2", "scales", "maptools", "sp", "maps", "grid", "car")
new.pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new.pkg)) {
  install.packages(new.pkg)
}
```

## Listing 3-2

Listing 3-2 uses an if statement to check whether a database exists, and then decides if it will use the download.file function to download a reputation database from the datadrivensecurity website to be saved in a subrepository of the working directory called data, as a file called reputation.data.

```{r 3-2}
avURL <- "http://datadrivensecurity.info/book/ch03/data/reputation.data"
avRep <- "data/reputation.data"
if (file.access(avRep)) {
  download.file(avURL, avRep)
}
```


## Listing 3-4

Listing 3-4 converts the # separated database into a dataframe using R, providing headers to the generated columns with the function colnames() and a vector of header strings.

```{r 3-4}
av <- read.csv(avRep, sep = "#", header = FALSE)
colnames(av) <- c("IP", "Reliability", "Risk", "Type", "Country", "Locale", "Coord", "X")
str(av)
```
```{r}
head(av)
```


## Listing 3-7

Listing 3-7 displays the 5 number summary developed by Tukey.  It is used to determine the range(min and max), and the first and third percentiles, along with the median and mean, of each specified column..

```{r 3-6}
summary(av$Reliability)
```
```{r}
summary(av$Risk)
```

## Listing 3-9

Listing 3-9 demonstrates how the table() function in R can count values of quantitative variables for a column in a dataframe.  Essentially, categorical data is aggregated and the count of each unique data is displayed.  It also shows the difference betwen table() and summary().  Summary organizes the malware qualitative data by aggregating each unique string and counting the number of times they appear

```{r 3-9}
table(av$Reliability)
```
```{r}
table(av$Risk)
```

```{r}
summary(av$Type, maxsum=10)
```

```{r}
summary(av$Country, maxsum=10)
```

##Listing 3-11

Listing 3-11 demonstrates the capabilities of ggplot2 library by creating a bar graph of the Country statistics in the dataset.

```{r 3-11}
library(ggplot2)
country.top20 <- names(summary(av$Country))[1:20]
gg <- ggplot(data=subset(av, Country %in% country.top20), 
             aes(x=reorder(Country, Country, length)))
gg <- gg + geom_bar(fill="#000099")
gg <- gg + labs(title="Country Counts", 
                x="Country", 
                y="Count")
gg <- gg + coord_flip()
gg <- gg + theme(panel.grid=element_blank(), 
                 panel.background=element_blank())
print(gg)
```

## Listing 3-12

Shows how the ggplot2 library can be used to create a bar graph of the number of each type of Categorical data in the Risk factor.

```{r}
gg <- ggplot(data=av, aes(x=Risk))
gg <- gg + geom_bar(fill="#000099")
gg <- gg + scale_x_discrete(limits=seq(max(av$Risk)))
gg <- gg + labs(title="'Risk' Counts", x="Risk Score", y="count")
gg <- gg + theme(panel.grid=element_blank(), panel.background=element_blank())
print(gg)
```

## Listing 3-13

Shows how the ggplot2 library can be used to create a bar graph of the number of each type of Categorical data in the Reliability factor.

```{r 3-13}
gg <- ggplot(data=av, aes(x=Reliability))
gg <- gg + geom_bar(fill="#000099")
gg <- gg + scale_x_discrete(limits=seq(max(av$Reliability)))
gg <- gg + labs(title="'Reliability' Counts", 
                x="Reliability Score", 
                y="Count")
gg <- gg + theme(panel.grid=element_blank(), 
                 panel.background=element_blank())
print(gg)
```


## Listing 3-17

TO look at the percentage of total malicious nodes contributed by the first 10 countries in the list, we divide each value by the number of rows in the dataframe.

```{r 3-17}
country10 <- summary(av$Country, maxsum=10)
country10.perc10 <- country10/nrow(av)
print(country10.perc10)
```

## Listing 3-19

A contingency table, which is a tabular view of the relationships between two variables, is used to determine which nodes to pay attention to when doing data-driven security analysis.  The xtabs is used to generate a matrix which represents quantity using size and colour.  This shows around where in the relationship the values in the dataset bias are concentrated.

```{r 3-19}
rr.tab <- xtabs(~Risk+Reliability, data=av)
ftable(rr.tab)
```
```{r}
library(lattice)
rr.df <- data.frame(table(av$Risk, av$Reliability))
colnames(rr.df) <- c("Risk", "Reliability", "Freq")
levelplot(Freq~Risk*Reliability, 
          data=rr.df, main="Risk ~ Reliability", 
          ylab="Reliability", 
          xlab="Risk", 
          shrink=c(0.5,1), 
          col.regions=colorRampPalette(c("#F5F5F5", "#01665E"))(20))
```

## Listing 3-21

Produces a matrix representing quantity with size and colour (levelplot) using random samples generated using random samples from the Risk (1:10) and Reliability (1:7) category (realization of the random process).  The randomness implies that it is unbiased, however, the process of selecting random samples may introduce its own bias, so multiple runs of the sample() function should be executed.  This visual is used to evaluate whether the real world data is due to chance or if there is meaning to the data. 

```{r}
set.seed(1492)
rel <- sample(1:7, 260000, replace=T)
rsk <- sample(1:10, 260000, replace=T)
tmp.df <- data.frame(table(factor(rsk), factor(rel)))
colnames(tmp.df) <- c("Risk", "Reliability", "Freq")
levelplot(Freq~Reliability*Risk, 
          data=tmp.df, 
          main="Risk ~ Reliability", 
          ylab="Reliability", 
          xlab="Risk", 
          shrink=c(0.5, 1), 
          col.regions=colorRampPalette(c("#F5F5F5", "#01665E"))(20))
```

## Listing 3-22

Compares each type of host to their Risk-Reliability measurement by creating a three-way contingency table.  Since Type can also be multiple types, the values are parsed so that those with the ';' character, indicating multiple types, is given their own category: "Multiples".

```{r}
av$simpletype <- as.character(av$Type)
av$simpletype[grep(';', av$simpletype)] <- "Multiples"
av$simpletype <- factor(av$simpletype)
rrt.df <- data.frame(table(av$Risk, av$Reliability, av$simpletype))
colnames(rrt.df) <- c("Risk", "Reliability", "simpletype", "Freq")
levelplot(Freq ~ Reliability * Risk | simpletype, 
          data=rrt.df, 
          main="Risk ~ Reliability | Type", 
          ylab="Risk", 
          xlab="Reliability", 
          shrink=c(0.5, 1), 
          col.regions=colorRampPalette(c("#F5F5F5", "#01665E"))(20))
```

## Listing 3-24

Omits the Scanning Host category from the three-way contingency table because the majority of entries are in that category and are generally low risk and reliability.

```{r}
rrt.df <- subset(rrt.df, simpletype != "Scanning Host")
levelplot(Freq ~ Reliability*Risk|simpletype, 
          data =rrt.df, 
          main="Risk ~ Reliabilty | Type", 
          ylab = "Risk", 
          xlab = "Reliability", shrink = c(0.5, 1), 
          col.regions = colorRampPalette(c("#F5F5F5","#01665E"))(20))
```

## Listing 3-26

Filters out Malware Domain from the three-way contingency graph since the majority is a risk and reliability around 2 and 3.  Also filters out Malware distribution since it does not seem to contribute any risk.

```{r 3-26}
rrt.df <- subset(rrt.df, !(simpletype %in% c("Malware distribution", "Malware Domain")))
sprintf("Count: %d; Percent: %2.1f%%", 
        sum(rrt.df$Freq), 
        100*sum(rrt.df$Freq)/nrow(av))
```
```{r}
levelplot(Freq ~ Reliability * Risk | simpletype, data=rrt.df, 
          main="Risk ~ Reliability | Type", ylab = "Risk", 
          xlab="Reliability", 
          shrink=c(0.5, 1), 
          col.regions=colorRampPalette(c("#F5F5F5", "#01665E"))(20))
```